\framework: Morphology-based Entity and Relational Entity Extraction Framework for Arabic

### 2018-04-20 tex/merf/TALMajor/tool.tex
done 25 pages

The set of tag types ${\cal T}$ contains tuples of the form $\langle l,f,d\rangle$ 
where $l$ is a text label with a descriptive name, 
$f$ is an MRE, and $d$ is a visualization legend 
with font and color information.

For each word $t_i\in T, 0\le i < M, M=|T|$.
\framework computes a Boolean value 
($\{\mathit{true}, \mathit{false}\}$)
for all MBFs. 
Then, it computes the set of MBF tags
$R_i=\{(t_i,tt)| tt=\langle l,f,d\rangle \wedge
f~\mathit{is~an~MBF} \wedge f(t_i)\} \subseteq T \times {\cal T}$
which tags a word $t_i$ with $\mathit{tt}$ 
iff the MBF $f$ associated with
tag type $\mathit{tt}$ is true for $t_i$. 
The MBF evaluation results in a sequence of tag sets 
$\langle R_0, R_1, \ldots, R_{n-1}\rangle$.
If a word $t_i$ has no tag type match, 
its tag set $R_i$ is by default the singleton $O=\{\mathit{NONE}\}$.
%and $t_i$ is referred to as a {\em null} word.

For each MRE, 
\framework generates its equivalent non-deterministic finite automaton (NFA) in the typical manner~\cite{sipser2006introduction}.
%Each MRE operation has its equivalent representation in an NFA. 
We support the upto operation ($f$\^{}$x$), which is not directly 
supported in~\cite{sipser2006introduction}, by 
%it is equivalent to $f?|ff|\dots|\underbrace{f}_{x \text{ times}}$ which has an NFA mapping. 
%we can expand it into a standard regular expression form, for example
expanding it into a regular expression form; for example 
$f$\^{}$3$ is equivalent to $f?|ff|fff$. 
Consider the example of Figure~\ref{fig:intromotiv} and the
corresponding expression $(P|N)\!+~O?~R~O^\wedge 2~(P|N|U)+$. 
Figure~\ref{fig:nfaEx} shows part of the corresponding NFA where
$q_8, q_9, \dots, q_{13}$ represent NFA states,
and edges are transitions based on MBF tags such as 
$P,$ and $N$.
Edges labeled with the empty string $\epsilon$ are non-deterministic.
%The expression uses operations $|$, $?$, $+$, and $\wedge$ to relate places {\tt P}, names of persons {\tt N}, relative positions {\tt R}, numerical terms {\tt U}, and other {\tt O}.

>>>>

The set of tag types ${\cal T}$ contains tuples of the form $\langle l,f,d\rangle$ 
where $l$ is a text label with a descriptive name, 
$f$ is an MRE, and $d$ is a visualization legend 
with font and color information.
\major{
For the example of Figure~\ref{fig:intromotiv},
$l$ is ``direction'',
$f$ is $(P|N)\!+~O?~R~O^\wedge 2~(P|N|U)+$,
and $d$ is italic. 
}

For each word $t_i\in T, 0\le i < |T|$.
\framework computes a Boolean value 
for all MBFs. 
\major{
For example, \RL{brj} matches MBF $P$. 
}
Then, it computes the set of MBF tags
$R_i=\{(t_i,tt)| tt=\langle l,f,d\rangle \wedge
f~\mathit{is~an~MBF} \wedge f(t_i)\} \subseteq T \times {\cal T}$
which tags a word $t_i$ with $\mathit{tt}$ 
iff the MBF $f$ associated with
tag type $\mathit{tt}$ is true for $t_i$. 
The MBF evaluation results in a sequence of tag sets 
$\langle R_0, R_1, \ldots, R_{n-1}\rangle$.
If a word $t_i$ has no tag type match, 
its tag set $R_i$ is by default the singleton $O=\{\mathit{NONE}\}$.
\major{
For example, $R$ for the text in Figure~\ref{f:overview} follows 
$\{ \{NONE\}, \{NONE\}, \{NONE\}, \{NONE\}, \{($\RL{brj}$,P)\}, \{($\RL{xlyfT}$,N)\},\ldots\}$. 
}

For each MRE, 
\framework generates its equivalent non-deterministic finite automaton (NFA) in the typical manner~\cite{sipser2006introduction}.
%Each MRE operation has its equivalent representation in an NFA. 
We support the upto operation ($f$\^{}$x$), which is not directly 
supported in~\cite{sipser2006introduction}, by 
%it is equivalent to $f?|ff|\dots|\underbrace{f}_{x \text{ times}}$ which has an NFA mapping. 
%we can expand it into a standard regular expression form, for example
expanding it into a regular expression form; for example 
$f$\^{}$3$ is equivalent to $f?|ff|fff$. 
Consider the example of Figure~\ref{fig:intromotiv} and the
corresponding expression $(P|N)\!+~O?~R~O^\wedge 2~(P|N|U)+$. 
\major{
Figure~\ref{fig:nfaEx} shows part of the corresponding NFA where
$q_8, q_9, \dots, q_{13}$ represent NFA states,
and edges are transitions based on MBF tags such as 
$P,$ and $N$.
Edges labeled with the empty string $\epsilon$ are non-deterministic.}
%The expression uses operations $|$, $?$, $+$, and $\wedge$ to relate places {\tt P}, names of persons {\tt N}, relative positions {\tt R}, numerical terms {\tt U}, and other {\tt O}.
